---
title: "FS20C3 - Datenjournalismus"
subtitle: "Data Analysis"
author: "Fernando Millan Villalobos"
date: "April 2020"
output:
  html_document:
    code_folding: show
    echo: TRUE
    warning: FALSE
    message: FALSE
    theme: paper
    df_print: kable
    toc: yes
    toc_depth: 4
    number_sections: yes
    toc_float: 
      collapsed: yes
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../", output_file = "index") })
---

```{r, echo=FALSE}
# CONFIG
user_name <- "FernandoMillanVillalobos" # your Git username (only needed if
# you want to deploy to GH pages)
project_name <- "fs20c3_datenjournalismus" # adapt!
package_date <- "2020-04-01" # date of the CRAN snapshot that
# the checkpoint package uses
r_version <- "4.0.0" # R-Version to use
options(Ncpus = 4) # use 4 cores for parallelized installation of packages
if (r_version != paste0(version$major, ".", version$minor)) {
  stop("ERROR: specified R version does not match currently used.")
}
```

# Notes

This report was generated on `r Sys.time()`. R version: `r paste0(version$major, ".", version$minor)` on `r version$platform`. For this report, CRAN packages as of `r package_date` were used.

## R-Script & data

The preprocessing and analysis of the data was conducted in the [R project for statistical computing](https://www.r-project.org/). The RMarkdown script used to generate this document and all the resulting data can be downloaded [under this link](http://`r user_name`.github.io/`r project_name`/). Through executing `main.Rmd`, the herein described process can be reproduced and this document can be generated. In the course of this, data from the folder `input` will be processed and results will be written to `output`. The html on-line version of the analysis can be accessed through this [link](https://`r user_name`.github.io/`r project_name`/).

## GitHub

The code for the herein described process can also be freely downloaded from [https://github.com/`r user_name`/`r project_name`](https://github.com/`r user_name`/`r project_name`). 

## Data description of output files

### `vietnam_bangladesh_wash.csv`

| year  | donor       | recipient  | amount   | code  | activity                              |
|-------|-------------|-----------------------------------------------------------------------|
| 2001  | ASDF        | Bangladesh | 55309088 | 14030 | Basic drinking water supply...        |
| 1997	| ASDB        | Viet Nam   | 3922870  | 14005 | Water Supply and Sanitation...        |
| 1981	| ASDB	      | Bangladesh | 1040441  | 14030 | Basic drinking water supply...        |

...

# Set up

```{r, echo=FALSE}
detach_all_packages <- function() {
  basic_packages_blank <-  c("stats",
                             "graphics",
                             "grDevices",
                             "utils",
                             "datasets",
                             "methods",
                             "base")
  basic_packages <- paste("package:", basic_packages_blank, sep = "")

  package_list <- search()[
    ifelse(unlist(gregexpr("package:", search())) == 1, TRUE, FALSE)]

  package_list <- setdiff(package_list, basic_packages)

  if (length(package_list) > 0)  for (package in package_list) {
    detach(package, character.only = TRUE, unload = TRUE)
    print(paste("package ", package, " detached", sep = ""))
  }
}

detach_all_packages()

# this allows multiple persons to use the same RMarkdown
# without adjusting the working directory by themselves all the time
source("scripts/csf.R")
path_to_wd <- csf() # if this - for some reason - does not work,
# replace with a hardcoded path, like so: "~/projects/rddj-template/analysis/"
if (is.null(path_to_wd) | !dir.exists(path_to_wd)) {
  print("WARNING: No working directory specified for current user")
} else {
  setwd(path_to_wd)
}

# suppress scientific notation
options(scipen = 999)

# unload global rstudioapi and knitr again to avoid conflicts with checkpoint
# this is only necessary if executed within RStudio
# outside of RStudio, namely in the knit.sh script, this causes RMarkdown
# rendering to fail, thus should not be executed there
if (Sys.getenv("RSTUDIO") == "1") {
  detach_all_packages()
}
```

## Define packages

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# from https://mran.revolutionanalytics.com/web/packages/\
# checkpoint/vignettes/using-checkpoint-with-knitr.html
# if you don't need a package, remove it from here (commenting not sufficient)
# tidyverse: see https://blog.rstudio.org/2016/09/15/tidyverse-1-0-0/
cat("
library(rstudioapi)
library(tidyverse) # ggplot2, dplyr, tidyr, readr, purrr, tibble, magrittr, readxl
library(scales) # scales for ggplot2
library(jsonlite) # json
library(lintr) # code linting
library(sf) # spatial data handling
library(rmarkdown)
library(cowplot) # theme
library(extrafont)", # fonts
file = "manifest.R")
```

## Install packages

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# if checkpoint is not yet installed, install it (for people using this
# system for the first time)
if (!require(checkpoint)) {
  if (!require(devtools)) {
    install.packages("devtools", repos = "http://cran.us.r-project.org")
    require(devtools)
  }
  devtools::install_github("RevolutionAnalytics/checkpoint",
                           ref = "v0.3.2", # could be adapted later,
                           # as of now (beginning of July 2017
                           # this is the current release on CRAN)
                           repos = "http://cran.us.r-project.org")
  require(checkpoint)
}
# nolint start
if (!dir.exists("~/.checkpoint")) {
  dir.create("~/.checkpoint")
}
# nolint end
# install packages for the specified CRAN snapshot date
checkpoint(snapshotDate = package_date,
           project = path_to_wd,
           verbose = T,
           scanForPackages = T,
           use.knitr = F,
           R.version = r_version)
rm(package_date)
```

## Load packages

```{r, echo=TRUE, message=FALSE, warning=FALSE}
source("manifest.R")
unlink("manifest.R")
sessionInfo()
```

## Load additional scripts

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# if you want to outsource logic to other script files, see README for 
# further information
# Load all visualizations functions as separate scripts
knitr::read_chunk("scripts/dviz.supp.R")
source("scripts/dviz.supp.R")
knitr::read_chunk("scripts/themes.R")
source("scripts/themes.R")
knitr::read_chunk("scripts/plot_grid.R")
source("scripts/plot_grid.R")
knitr::read_chunk("scripts/align_legend.R")
source("scripts/align_legend.R")
knitr::read_chunk("scripts/label_log10.R")
source("scripts/label_log10.R")
```

# Introduction
This notebook is thought as a working document for the development and documentation of the [FS20C3 - Datenjournalismus project](https://ds-spaces.technik.fhnw.ch/app/uploads/sites/45/2019/09/hs19c3datenjournalsimus.pdf)).

[AidData](https://www.aiddata.org/) is a research lab at William & Mary's Global Research Institute. AidData’s interdisciplinary team focus on research and analysis of development investments from more than 90 grants and contracts worth more than $50 million over the last 16 years. Taking as starting point the [data](https://www.aiddata.org/data/aiddata-core-research-release-level-1-3-1) provided by them and after a previous research about media use of the information gathered and made available to the public by this organization, we´ve developed a new data-driven story. In order to do that, data from the core dataset were selected, extracted and analysed to built our written and visual narrative.

The data-driven story we want to tell with the help of the data provided by AidData, is about the succesfull and quick development of a national water supply and sanitation net that Vietnam has accomplished in the past two decades. This huge progress has been possible thanks, among other reasons, to a broad international cooperation. In this notebook we´ve digged into the available data to bring into the light the main protagonists of this collaborative achievement and try to picture the timeline that makes it a today reality.

With this in mind, we´ve looked for the 3-digit sector classification based on OECD purpose codes that comes in [the documentation of the data](http://docs.aiddata.org/ad4/files/aiddata_coding_scheme_0.pdf) and selected the activities that are related to our main goal.

# Data Importing
The core dataset provided by AidData can be found [here](https://www.aiddata.org/data/aiddata-core-research-release-level-1-3-1). This dataset tracks known international development flows from bilateral and multilateral aid organizations from 1947 to 2013 in a flat-table structure. Information available are served as a granular data from more level of details to a general overview. After getting a first impression of the different datasets that encompass the whole data core and reading the [documentation](https://www.aiddata.org/methods/sdg-coding), we end that the information we need for our story is located in the "thin" version (more than 1.500.000 entries).

```{r importing data}

# Importing whole dataset
df_full <- read_csv("input/AidDataCoreThin_ResearchRelease_Level1_v3.1.csv")
head(df_full)

```

# Data Pre-Processing
In this step we are going to transform our imported data in a way that allows us to apply further analysis. Typically, this process requires some data wrangling, data transformation and the creation of new variables.

```{r selecting filtering renaming}

# The 3-digit sector classification based on OECD purpose codes tells us that
# codes beginning with 140 corresponds to “Water Supply and Sanitation”.
# Among the different activities in this category:
# - Water Supply and Sanitation, combination of purposes (14000)
# - Water Supply and Sanitation, purpose unspecified or does not fit under any other applicable codes (14005)
# - Basic drinking water supply and basic sanitation (14030) 
# - Waste management/disposal (14050)
# - Education and training in water supply and sanitation (14081) 

# Select columns and rows we are interesting in with values different as na´s.
codes <- c(14000, 14005, 14030, 14050, 14081)
(head(df1 <- df_full %>%
  select(3:8) %>%
  filter((recipient == "Viet Nam" | recipient == "Bangladesh") & coalesced_purpose_code %in% codes) %>% # Subjects of our analysis are Viet Nam and Bangladesh
  na.omit()))

# Select columns and rows of our operational dataset
(head(df <- df1 %>%
  rename(amount = commitment_amount_usd_constant, code = coalesced_purpose_code, activity = coalesced_purpose_name)))

df$donor <- df$donor %>% 
  as_factor() %>% # convert variable donor as factor
  recode_factor(
    "Asian Development Bank (ASDB)" = "ASDB",
    "Asian Development Fund (ASDF)" = "ASDF",
    "Asian Development Bank (AsDB Special Funds)" = "AsDB",
    "World Bank - International Development Association (IDA)" = "IDA",
    "United Nations Children`s Fund (UNICEF)" = "UNICEF",
    "OPEC Fund for International Development (OFID)" = "OFID",
    "European Communities (EC)" = "EC",
    "International Fund for Agricultural Development (IFAD)" = "IFAD",
    "Bill & Melinda Gates Foundation" = "B&M",
    "Global Environment Facility (GEF)" = "GEF"
  )

df$code <- df$code %>% 
  as_factor() # convert variable code as factor

df$activity <- df$activity %>% 
  as_factor() # convert variable activity as factor

df_vietnam <- df %>% 
  filter(recipient == "Viet Nam")

df_bangladesh <- df %>% 
  filter(recipient == "Bangladesh")

# Export our operational dataset as filtered csv file
# write_csv(df, "output/vietnam_bangladesh_wash.csv")

```

# Exploratory Data Analysis (EDA)
## Data Summarization
Now we want to describe some important properties of the distribution of the values across the observations in our datasets, an overview of the key properties of the data.

```{r eda summarising donors}

# How many different donors
(distinct_donors <- unique(df$donor) %>% 
  n_distinct())

# Times and frequency donors contribution
(donors_contrib <- df %>% 
  group_by(donor) %>% 
  summarize(
    count = n()
  ) %>% 
  arrange(desc(count)) %>%
  mutate(freq = count / sum(count),
         pct = round(freq * 100, 1)) %>% 
  select(donor, count, pct))

```

From our primary data exploration we can see that there´re a total of __`r distinct_donors`__. Donors that have contributed more frequently are UNICEF (__`r donors_contrib$pct[1]`%__) , Germany (__`r donors_contrib$pct[2]`%__) and Japan (__`r donors_contrib$pct[3]`%__). These three main donors represent together up to __`r donors_contrib$pct[1] + donors_contrib$pct[2] + donors_contrib$pct[3]`%__ from all contributions. On the other hand, donors like New Zealand (__`r donors_contrib$pct[31]`%__), United Arab Emirates (__`r donors_contrib$pct[30]`%__) and IFAD (__`r donors_contrib$pct[29]`%__) are located at the bottom of the distribution with an all together total percentage of less than __1%__.

```{r data summarising amount}

# Finding the most common values of our quantitative variable in our sample.
select(df, amount) %>%
  summarise_each(list(max = max, min = min, mean = mean, median = median, var = var, sd = sd, IQR = IQR))

# Donors amount ranking distribution
(donors_ranking_amount <- df %>% 
  group_by(donor) %>% 
  summarize(
    total_amount = sum(amount),
    mean = mean(amount),
    median = median(amount),
    max = max(amount),
    min = min(amount)
  ) %>% 
  arrange(desc(total_amount)) %>% 
  mutate(freq = total_amount / sum(total_amount),
         pct = round(freq * 100, 3)) %>% 
  select(donor, total_amount, pct, mean, median, max, min))

# Total amount given for all donors to Vietnam from 1992 to 2013
(total_amount_donors <- sum(donors_ranking_amount$total_amount))

```

Looking at our first preliminary summarized results for the variable amount, we can see that surprisingly, among the more frequent donors, just Japan (__`r donors_ranking_amount$pct[1]`%__) stays in the podium as the most generous contributors. The other two, Germany (__`r donors_ranking_amount$pct[7]`%__) and UNICEF (__`r donors_ranking_amount$pct[18]`%__), are down (7º Germany) or far down (18º UNICEF) the list. Just as we saw before, together the three main donors (1º Japan, 2º IDA and 3º ASDB) regarding the amount of money they´ve given in the past two decades to Vietnam and Bangladesh, represents a vast proportion of the total amount of money given adding up to __`r donors_ranking_amount$pct[1] + donors_ranking_amount$pct[2] + donors_ranking_amount$pct[3]`%__. Another thing that catched our eye was the huge difference between max and min values. Let´s take our main contributor, Japan, as example. The maximum amount of money that that country has given to our both analysed countries was __`r donors_ranking_amount$max[1]` USD__, while the minimum was just __`r donors_ranking_amount$min[1]` USD__. A __`r donors_ranking_amount$max[1] * donors_ranking_amount$min[1] / 100`%__ more!

The total amount both countries has received from the 31 donors during the period that goes from 1992 to 2013 (dataset) was __`r total_amount_donors` USD__.

## Data Visualization
We´re going to explore and gain insights from our data using one of the most outstanding ability of human beings: capturing visual patterns. Let´s explore distributions of key variables (donor, amount and activity), associations among them and time development in our dataset. 

### Donors Distribution

```{r donors distribution}

# Distinct donors


# p <- ggplot(subset(df, amount > 10000000), aes(x = donor, y = amount)) +
#   geom_col() +
#   labs(title = "Mass distribution in Dataset 01") +
#   coord_cartesian(clip = "off") +
#   theme_dviz_vgrid() +
#   theme(
#     axis.line.x = element_blank(),
#     plot.margin = margin(3, 7, 3, 1.5)
#   ) +
#   coord_flip()
# p

```


## Linting

The code in this RMarkdown is linted with the [lintr package](https://github.com/jimhester/lintr), which is based on the  [tidyverse style guide](http://style.tidyverse.org/). 

```{r echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
lintr::lint("main.Rmd", linters =
              lintr::with_defaults(
                commented_code_linter = NULL,
                trailing_whitespace_linter = NULL
                )
            )
# if you have additional scripts and want them to be linted too, add them here
# lintr::lint("scripts/my_script.R")
```
